---
title: "Introduction to Forecasting"
subtitle: "Predicting Salmon Returns"  
author: 
#  - "Mark Scheuerell"
  - "WA-BC AFS Meeting<br>Spokane, WA<br><br>29 April 2024"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "my-theme.css"]
    nature:
      slideNumberFormat: ""
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      # countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=8, fig.height=4, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)

## set random seed
set.seed(666)

## load {icons} for font awesome
library(icons)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo(
  primary_color = "#FFFFFF",
  secondary_color = "#23395b",
  # secondary_color = "#FF961C",
  title_slide_text_color = "#FFFFFF",
  title_slide_background_color = "#23395b",
  colors = c(
    red = "#f34213",
    purple = "#844870",
    orange = "#FF961C",
    green = "#339933",
    white = "#FFFFFF",
    blue = "#488fdf",
    aqua = "#80cdc1",
    gray = "#808080",
    lightgray = "#bdbdbd"
),
  header_font_google = google_font("Roboto Condensed", "400", "400i", "700"),
  text_font_google = google_font("Cabin", "400", "400i", "700"),
)
```


# Forecasting life cycle

.pull-left[
### .blue[Assess & understand objectives]
]


---

# Forecasting life cycle

.pull-left[
### .blue[Assess & understand objectives]

### .blue[Gather information]
]


---

# Forecasting life cycle

.pull-left[
### .blue[Assess & understand objectives]

### .blue[Gather information]

### .blue[Decide on modeling framework]
]

---

# Forecasting life cycle

.pull-left[
### .blue[Assess & understand objectives]

### .blue[Gather information]

### .blue[Decide on modeling framework]

### .blue[Choose fitting & forecast horizons]
]

---

# Forecasting life cycle

.pull-left[
### .blue[Assess & understand objectives]

### .blue[Gather information]

### .blue[Decide on modeling framework]

### .blue[Choose fitting & forecast horizons]
]

.pull-right[
### .blue[Fit model(s) & generate forecast(s)]
]


---

# Forecasting life cycle

.pull-left[
### .blue[Assess & understand objectives]

### .blue[Gather information]

### .blue[Decide on modeling framework]

### .blue[Choose fitting & forecast horizons]
]

.pull-right[
### .blue[Fit model(s) & generate forecast(s)]

### .blue[Evaluate skill]
]


---

# Forecasting life cycle

.pull-left[
### .blue[Assess & understand objectives]

### .blue[Gather information]

### .blue[Decide on modeling framework]

### .blue[Choose fitting & forecast horizons]
]

.pull-right[
### .blue[Fit model(s) & generate forecast(s)]

### .blue[Evaluate skill]

### .blue[Report results]

]


---

# Forecasting life cycle

.pull-left[
### .blue[Assess & understand objectives]

### .blue[Gather information]

### .blue[Decide on modeling framework]

### .blue[Choose fitting & forecast horizons]
]

.pull-right[
### .blue[Fit model(s) & generate forecast(s)]

### .blue[Evaluate skill]

### .blue[Report results]

### .blue[Rinse & repeat]
]


---

# Assess & understand objectives

## .blue[Who is involved?]

--

## .blue[Harvest vs conservation?]

--

## .blue[Escapement goal vs exploitation rate?]

--

## .blue[Pre- vs in-season?]


---

# Gather information

## .blue[Who holds the data?]

--

## .blue[Harvest + escapement?]

--

## .blue[Age composition?]

--

## .blue[Smolts?]

--

## .blue[Environmental covariates?]


---

# Decide on modeling framework

## .blue[Sibling regression vs environmental? Both?]

--

## .blue[Observation & process errors? (state-space)]

--

## .blue[Counts? Age-comp? Tags? (integrated population model)]


---

# Choose fitting & forecast horizons

```{r forecast_horizon, out.width = "80%", out.height = "80%"}



```



---

# Fit model(s) & generate forecast(s)

## .blue[Bayesian? Priors?]

--

## .blue[Software?]


---

# Evaluate skill

## .blue[Accuracy?]

--

## .blue[Precision?]

--

## .blue[Formal model comparisons?]


---

# Report results

## .blue[Open and reproducible?]




